ethan_analysis

### 2022-12-09 austins_temp_powerpoint.rmd

---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
knitr::opts_chunk$set(fig.width=12, fig.height=8) 
```

# Breast Cancer Diagnosis Classification Analysis

>>>>

---
```{r setup, include=FALSE}
# Import packages
library(tidyverse)
library(ggplot2)
library(GGally)
library(knitr)
library(boot)
library(caret)
library(randomForest)
library(e1071)
library(class)
knitr::opts_chunk$set(fig.width=14, fig.height=8)
set.seed(420)
```

```{r load_data, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
# Load the data
data <- read.csv("breast_cancer_diagnostic_data.csv") # Full Data
training <- read.csv("training.csv") # Training Data
test <- read.csv("test.csv")  # Test Data
# Data Cleaning
data <- data[,-1] # Remove ID column
test <- test[,-1] # Remove ID column
training <- training[,-1] # Remove ID column
# Convert diagnosis to factor
data$diagnosis <- as.factor(data$diagnosis)
test$diagnosis <- as.factor(test$diagnosis)
training$diagnosis <- as.factor(training$diagnosis)
```

## Background

- Many people are diagnosed with breast cancer each year
- Breast Cancer Diagnosis Data
- Predict Diagnosis of Tumor based on features (Malignant or Benign)
- *Modivation:* Reduce Type II Error (False-Negatives)
  - Feature and Model Selection

# Models

```{r train_setup, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
train_control <- trainControl(method = "repeatedcv", summaryFunction = defaultSummary, classProbs = TRUE, number = 10, repeats = 10)
```

# KNN Model

## Tuning KNN

```{r knn_tuning, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
type_two_error_knn <- data.frame(k = seq(1, 100, 1), error = NA, typetwoerror = NA)

for (i in 1:nrow(type_two_error_knn)) {
    knn <- train(diagnosis ~ ., data = training, method = "knn", trControl = train_control, metric = "Accuracy", tuneGrid = data.frame(k = type_two_error_knn$k[i]))
    knn_pred <- predict(knn, test)
    knn_confusion <- confusionMatrix(knn_pred, test$diagnosis, positive = "M")
    type_two_error_knn$typetwoerror[i] <- knn_confusion$table[1, 2]/sum(knn_confusion$table)
    type_two_error_knn$error[i] <- (knn_confusion$table[1, 2] + knn_confusion$table[2, 1])/sum(knn_confusion$table)
}
print(which.min(type_two_error_knn$typetwoerror))
knn1_k <- which.min(type_two_error_knn$typetwoerror)
knn_tte <- min(type_two_error_knn$typetwoerror)
```

```{r knn_tuning_plot, echo=FALSE, message=FALSE, warning=FALSE}
ggplot() + geom_line(aes(x=type_two_error_knn$k, y=type_two_error_knn$error), size=1) + geom_line(aes(x=type_two_error_knn$k, y=type_two_error_knn$typetwoerror), size=1.5, color="#c82798") + labs(title = "KNN Type I and Overall Error", x = "K", y = "Error")
```

## Tuned KNN

```{r tuned_knn, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
set.seed(420)
tuned_knn <- train(diagnosis ~ ., data = training, method = "knn", trControl = train_control, metric = "Accuracy", tuneGrid = data.frame(k = which.min(type_two_error_knn$typetwoerror)))
knn_pred <- predict(tuned_knn, test)
knn_confusion <- confusionMatrix(knn_pred, test$diagnosis, positive = "M")
```

```{r tuned_knn_plot, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data.frame(knn_confusion$table), aes(x = Prediction, y = Reference, fill = Freq)) +
  geom_tile() +
  geom_text(aes(label = Freq), size = 12) +
  scale_fill_gradient(low = "white", high = "#c82798") +
  labs(title = "KNN Confusion Matrix", x = "Prediction", y = "Reference") +
  theme_minimal() +
  theme(axis.text = element_text(size = 20), axis.title = element_text(size = 30), plot.title = element_text(size = 30)) +
  annotate("path",
   x=1+.25*cos(seq(0,2*pi,length.out=100)),
   y=2+.25*sin(seq(0,2*pi,length.out=100)), size=1.5, linetype="dashed", color="#c82798")
```

## Tuned KNN Parameters

- Using all features
- 10-fold cross validation repeated 10 times
- $k=$ `r knn1_k` neighbors
- Type II Error: `r knn_tte`
- Accuracy: `r knn_confusion$overall[1]`

# Random Forrest

## Tuning Random Forrest

```{r rf_tuning, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
type_two_error_rf <- data.frame(ntree = seq(1, 100, 10), error = NA, typetwoerror = NA)

for (i in 1:nrow(type_two_error_rf)) {
    rf <- train(diagnosis ~ ., data = training, method = "rf", trControl = train_control, metric = "Accuracy", ntree = type_two_error_rf$ntree[i])
    rf_pred <- predict(rf, test)
    rf_confusion <- confusionMatrix(rf_pred, test$diagnosis, positive = "M")
    type_two_error_rf$typetwoerror[i] <- rf_confusion$table[1, 2]/sum(rf_confusion$table)
    type_two_error_rf$error[i] <- (rf_confusion$table[1, 2] + rf_confusion$table[2, 1])/sum(rf_confusion$table)
}
print(which.min(type_two_error_rf$typetwoerror))
rf1_ntree <- type_two_error_rf$ntree[which.min(type_two_error_rf$typetwoerror)]
rf_tte <- min(type_two_error_rf$typetwoerror)
```

```{r rf_tuning_plot, echo=FALSE, message=FALSE, warning=FALSE}
ggplot() + geom_line(aes(x=type_two_error_rf$ntree, y=type_two_error_rf$error), size=1) + geom_line(aes(x=type_two_error_rf$ntree, y=type_two_error_rf$typetwoerror), size=1.5, color="#c82798") + labs(title = "Random Forrest Type I and Overall Error", x = "Number of Trees", y = "Error")
```

## Tuned Random Forrest

```{r tuned_rf, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
tuned_rf <- train(diagnosis ~ ., data = training, method = "rf", trControl = train_control, metric = "ROC", ntree = type_two_error_rf$ntree[which.min(type_two_error_rf$typetwoerror)])
rf_pred_tuned <- predict(tuned_rf, test)
rf_confusion <- confusionMatrix(rf_pred_tuned, test$diagnosis, positive = "M")
```

```{r tuned_rf_plot, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data.frame(rf_confusion$table), aes(x = Prediction, y = Reference, fill = Freq)) +
  geom_tile() +
  geom_text(aes(label = Freq), size = 12) +
  scale_fill_gradient(low = "white", high = "#c82798") +
  labs(title = "Random Forrest Confusion Matrix", x = "Prediction", y = "Reference") +
  theme_minimal() +
  theme(axis.text = element_text(size = 20), axis.title = element_text(size = 30), plot.title = element_text(size = 30)) +
  annotate("path",
   x=1+.25*cos(seq(0,2*pi,length.out=100)),
   y=2+.25*sin(seq(0,2*pi,length.out=100)), size=1.5, linetype="dashed", color="#c82798")
```

## Tuned Random Forrest Parameters

- Using all features
- 10-fold cross validation repeated 10 times
- $n=$ `r rf1_ntree` trees
- Type II Error: `r rf_tte`
- Accuracy: `r rf_confusion$overall[1]`

## Radial Support Vector Machine

```{r austin_svm, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
#Radial Support Vector Machine
svm <- train(diagnosis ~ ., data = training, method = "svmRadial", trControl = train_control, metric="ROC")
svm_pred <- predict(svm, test)
svm_confusion <- confusionMatrix(svm_pred, test$diagnosis, positive="M")

svm_tte <- svm_confusion$table[1, 2]/sum(svm_confusion$table)
```

```{r austin_svm_plot, echo=FALSE, message=FALSE, warning=FALSE}

ggplot(data.frame(svm_confusion$table), aes(x = Prediction, y = Reference, fill = Freq)) +
  geom_tile() +
  geom_text(aes(label = Freq), size = 12) +
  scale_fill_gradient(low = "white", high = "#c82798") +
  labs(title = "Radial SVM Confusion Matrix", x = "Prediction", y = "Reference") +
  theme_minimal() +
  theme(axis.text = element_text(size = 20), axis.title = element_text(size = 30), plot.title = element_text(size = 30)) +
  annotate("path",
   x=1+.25*cos(seq(0,2*pi,length.out=100)),
   y=2+.25*sin(seq(0,2*pi,length.out=100)), size=1.5, linetype="dashed", color="#c82798")
```

## Radial Support Vector Machine Parameters

- Using all features
- 10-fold cross validation repeated 10 times
- Type II Error: `r svm_tte`
- Accuracy: `r svm_confusion$overall[1]`

## Tuned Radial Support Vector Machine

```{r austin_svm2, results=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
tuned_svm <- tune.svm(diagnosis ~., data = training, gamma = 10^(-5:-1), cost = 10^(-3:1))
svm2 <- svm(diagnosis ~., data = training, gamma = 0.01, cost = 10, type="C-classification", kernel="radial")
svm2_pred <- predict(svm2, test)
svm2_confusion <- confusionMatrix(svm2_pred, test$diagnosis, positive="M")

svm2_tte <- svm2_confusion$table[1, 2]/sum(svm2_confusion$table)

```

```{r austin_svm_plot2, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data.frame(svm2_confusion$table), aes(x = Prediction, y = Reference, fill = Freq)) +
  geom_tile() +
  geom_text(aes(label = Freq), size = 12) +
  scale_fill_gradient(low = "white", high = "#c82798") +
  labs(title = "Tuned Radial SVM Confusion Matrix", x = "Prediction", y = "Reference") +
  theme_minimal() +
  theme(axis.text = element_text(size = 20), axis.title = element_text(size = 30), plot.title = element_text(size = 30)) +
  annotate("path",
   x=1+.25*cos(seq(0,2*pi,length.out=100)),
   y=2+.25*sin(seq(0,2*pi,length.out=100)), size=1.5, linetype="dashed", color="#c82798")
```

## Tuned Radial Support Vector Machine Parameters

- Using all features
- 10-fold cross validation repeated 10 times
- Type II Error: `r svm2_tte`
- Accuracy: `r svm2_confusion$overall[1]`

### 2022-12-09 ethan_analysis.Rmd

```

@@

```{r}
#Test
am_lm <- lm(area_mean ~ ., data = trainTot)
pred_amtest <- predict(am_lm, testTot)
amtest_log <- postResample(pred_amtest, testTot$area_mean)
amtest_log
#summary(am_lm)
```

```{r}
#Train
pred_amtrain<- predict(am_lm, trainTot)
amtrain_log <- postResample(pred_amtrain, trainTot$area_mean)
amtrain_log
```

@@

@@

predictions_test <- predict(lasso_model, s = lambda_best, newx = x_test)
aml_test <- eval_results(y_test, predictions_test, testTot)
```

@@

predictions_test <- predict(lasso_model, s = lambda_best, newx = x_test)
awl_test <- eval_results(y_test, predictions_test, testTot)
```

>>>>

head(dataTot)
```

@@

```{r}
#Train
am_lm <- lm(area_mean ~ ., data = trainTot)
pred_amtrain<- predict(am_lm, trainTot)
amtrain_log <- postResample(pred_amtrain, trainTot$area_mean)
amtrain_log
```
```{r}
#Test
pred_amtest <- predict(am_lm, testTot)
amtest_log <- postResample(pred_amtest, testTot$area_mean)
amtest_log
#summary(am_lm)
```

@@

  # Model performance metrics
data.frame(
  RMSE = RMSE,
  Rsquared = R_square)

@@

predictions_test <- predict(lasso_model, s = lambda_best, newx = x_test)
aml_test <- eval_results(y_test, predictions_test, testTot)

print(aml_train)
print(aml_test)
```

```{r}
as_data_frame(rbind(amtrain_log,
      amr_train,
      aml_train,
      amtest_log,
      amr_test,
      aml_test)) %>%
    mutate(model = c('Linear Train Error',  'Ridge Train Error', 'Lasso Train Error',  'Linear Test Error','Ridge Test Error',  'Lasso Test Error')) %>%
    select(model, RMSE, Rsquared)
```

@@

predictions_test <- predict(lasso_model, s = lambda_best, newx = x_test)
awl_test <- eval_results(y_test, predictions_test, testTot)

print(awl_train)
print(awl_test)
```

```{r}
as_data_frame(rbind(awtrain_log,
      awr_train,
      awl_train,
      awtest_log,
      awr_test,
      awl_test)) %>%
    mutate(model = c('Linear Train Error',  'Ridge Train Error', 'Lasso Train Error',  'Linear Test Error','Ridge Test Error',  'Lasso Test Error')) %>%
    select(model, RMSE, Rsquared)
```
--------------------------------------------------------------------------------
## concave.points_mean

```{r}
#Train
cvptsm_lm <- lm(concave.points_mean ~ ., data = trainTot)
pred_cvptsm_train<- predict(cvptsm_lm, trainTot)
cvptsm_train_log <- postResample(pred_cvptsm_train, trainTot$concave.points_mean)
cvptsm_train_log
```
```{r}
#Test
pred_cvptsm_test <- predict(cvptsm_lm, testTot)
cvptsm_test_log <- postResample(pred_cvptsm_test, testTot$concave.points_mean)
cvptsm_test_log
#summary(am_lm)
```

# ridge regression
```{r}
dummy <- dummyVars(concave.points_mean~., data = dataTot[,-1])
x = as.matrix(predict(dummy, newdata = trainTot[,-1]))
y_train = trainTot$concave.points_mean

x_test = as.matrix(predict(dummy, newdata = testTot[,-1]))
y_test = testTot$concave.points_mean

# Prediction and evaluation on train data
predictions_train <- predict(ridge_reg, s = optimal_lambda, newx = x)
cvptsm_train <- eval_results(y_train, predictions_train, trainTot)

# Prediction and evaluation on test data
predictions_test <- predict(ridge_reg, s = optimal_lambda, newx = x_test)
cvptsm_test <- eval_results(y_test, predictions_test, testTot)

print(cvptsm_train)
print(cvptsm_test)
```

predictions_train <- predict(lasso_model, s = lambda_best, newx = x)
cvptsm_train <- eval_results(y_train, predictions_train, trainTot)

predictions_test <- predict(lasso_model, s = lambda_best, newx = x_test)
cvptsm_test <- eval_results(y_test, predictions_test, testTot)

print(cvptsm_train)
print(cvptsm_test)
```

## Radius Worst

```{r}
#Train
rw_lm <- lm(radius_worst ~ ., data = trainTot)
pred_rw_train<- predict(rw_lm, trainTot)
rw_train_log <- postResample(pred_rw_train, trainTot$radius_worst)
rw_train_log
```
```{r}
#Test
pred_rw_test <- predict(rw_lm, testTot)
rw_test_log <- postResample(pred_rw_test, testTot$radius_worst)
rw_test_log
#summary(am_lm)
```

# ridge regression
```{r}
dummy <- dummyVars(radius_worst~., data = dataTot[,-1])
x = as.matrix(predict(dummy, newdata = trainTot[,-1]))
y_train = trainTot$radius_worst

x_test = as.matrix(predict(dummy, newdata = testTot[,-1]))
y_test = testTot$radius_worst

# Prediction and evaluation on train data
predictions_train <- predict(ridge_reg, s = optimal_lambda, newx = x)
rw_train <- eval_results(y_train, predictions_train, trainTot)

# Prediction and evaluation on test data
predictions_test <- predict(ridge_reg, s = optimal_lambda, newx = x_test)
rw_test <- eval_results(y_test, predictions_test, testTot)

print(rw_train)
print(rw_test)
```

predictions_train <- predict(lasso_model, s = lambda_best, newx = x)
rw_train <- eval_results(y_train, predictions_train, trainTot)

predictions_test <- predict(lasso_model, s = lambda_best, newx = x_test)
rw_test <- eval_results(y_test, predictions_test, testTot)

print(rw_train)
print(rw_test)
```
